---
title: "R vs  Python: Different similarities and similar differences"
author: "Tinniam V Ganesh"
date: "May 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R versus Python 
As to whic language is mosr suitd for Datascience and Analysis, there has been a lot of debate with
people tending to get quitre dear to them. This post will look at some of the similarities and differences
To a large extent bothR an Python do the same thing in just slightly different ways and syntaxes. The ease or the difficulty in the R/Pythin construct is in the 'eyes of the beholder nay, progeammer' O would say.\

- R has the type vector which in Python is closely related to a list
- To know the type of the variable we use the following for R and Python
- R & Python use length & len respectively to get the length of a vector and list respectively

# R Vector vs Python List
```{r}
# R vectors
a<-c(4,5,1,3,4,5)
print(a[3])
print(a[3:4]) # R does not always need the explicit print. 
#R type of variable
print(class(a))
# Length of a
print(length(a))
```

```{python}
# Python lists
a=[4,5,1,3,4,5] # 
print(a[2]) # Some python IDEs require the explicit print
print(a[2:5])
print(type(a))
# Length of a
print(len(a))
```

Python also has certain other data types like the tuple, dictionary etc as shown below. R does not have as many of the data types, nevertheless we can do everything that Python does in R

```{python}
# Python tuple
b = (4,5,7,8)
print(b)


#Python dictionary
c={'name':'Ganesh','age':54,'Work':'Professional'}
print(c)
#Print type of variable c

```


###2.Variable type
To know the type of the variable in R we use 'class'

```{r}
#R - Type of variable
a<-c(4,5,1,3,4,5)
print(class(a))

```

```{python}
#Python - Print type of tuple a
a=[4,5,1,3,4,5]
print(type(a))
b=(4,3,"the",2)
print(type(b))
```

###3 Length 
To know length in R, use length()
```{r}
#R - Length of vector
# Length of a
a<-c(4,5,1,3,4,5)
print(length(a))
```

To know the length of a list,tuple or dict we can use
```{python}
# Python - Length of list , tuple etc
# Length of a
a=[4,5,1,3,4,5]
print(len(a))
# Length of b
b = (4,5,7,8)
print(len(b))

```


###4. Accessing help

To access help in R we use the '?' or the 'help' function 
```{r}
#R - Help - To be done in R console or RStudio
#?sapply
#help(sapply)
```

Help in python on any topic involves
```{python}
#Python help - This can be done on a (I)Python console
#help(len)
#?len
```

###5. Subsetting
The key difference between R and Python with regards to subsetting is that in R the index starts at 1. In Python it starts at 0, much like C,C++ or Java
To subset a vector in R we use
```{r}
#R - Subset
a<-c(4,5,1,3,4,8,12,18,1)
print(a[3])

# To print a range or a slice. Print from the 3rd to the 5th element
print(a[3:6])
```

```{python}
#Python - Subset
a=[4,5,1,3,4,8,12,18,1]
# Print the 4th element (starts from 0)
print(a[3])

# Print a slice from 4 to 6th element
print(a[3:6])
```

###6. Operations on vectors in R and operation on lists in Python
In R we can do many operations on vectors for e.g. element by element addition, subtraction,
exponentation,product etc. as show
```{r}
#R - Operations on vectors
a<- c(5,2,3,1,7)
b<- c(1,5,4,6,8)

#Element wise Addition
print(a+b)
#Element wise subtraction
print(a-b)
#Element wise product
print(a*b)
# Exponentiating the elements of a vector
print(a^2)
```

In Python to do this on lists we need to use the 'map' and the 'lambda' function as follows
```{python}
# Python - Operations on list
a =[5,2,3,1,7]
b =[1,5,4,6,8]

#Element wise addition with map & lambda
print(map(lambda x,y: x+y,a,b))
#Element wise subtraction
print(map(lambda x,y: x-y,a,b))
#Element wise product
print(map(lambda x,y: x*y,a,b))
# Exponentiating the elements of a list
print(map(lambda x: x**2,a))


```

However if we create ndarrays from lists then we can do the element wise addition,subtraction,product,
etc. like R. Numpy is really a powerful module with many, many functions for matrix manipulations
```{python}
import numpy as np
a =[5,2,3,1,7]
b =[1,5,4,6,8]
a=np.array(a)
b=np.array(b)
#Element wise addition
print(a+b)
#Element wise subtraction
print(a-b)
#Element wise product
print(a*b)
# Exponentiating the elements of a list
print(a**2)

```

###7 Getting the index of element
To determine the index of an element which satisifies a specific logical condition in R use 'which'. In the code below the index of  element which is equal to 1 is 4
```{r}
# R - Which
a<- c(5,2,3,1,7)
print(which(a == 1))
```

In Python array we can use np.where to get the same effect. The index will be 3 as the index starts from 0
```{python}
# Python - np.where
import numpy as np
a =[5,2,3,1,7]
a=np.array(a)
print(np.where(a==1))
```

###8. Data frames
R, by default comes with a set of datasets. There are some datasets which come with the SkiKit- Learn package

```{r}
# R 
# To check built datasets use
#data() - In R console or in R Studio
iris

```


```{python}
#Python
import sklearn as sklearn
from sklearn import datasets
iris = datasets.load_iris()
iris
```

